Temperature Measurements in Optical Tweezer Experiments

### 2017-01-12 doc/draft1.tex
more draft, more graphs

\subsection{Velocity Verlet}
The simulation is taking place in a NVE ensemble from a thermodynamic point of view. This means that the particle number $N$, the volume $V$ and the
total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
they have been drawn.\\
The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
of all the atoms in the system have to be added to get the resulting velocity vector for the whole system that is thought to be acting on the center
of mass. This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
velcocity vector of every atom in the system. 
\begin{eqnarray}
    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
\end{eqnarray}
The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
The angular momentum can be defined via the moment of inertia $\mathbf{I}$
\begin{eqnarray}
    \mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
    \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
\end{eqnarray}
where $\boldsymbol{\omega}$ is the angular velocity of the system. 
\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. If the system should have a
specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
\begin{eqnarray}
    E &=& \sum_{i=1}^N \mathbf{v}_i^2\\
    \lambda &=& \sqrt{\frac{3(N-1)T}{E}}\\
    \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i 
\end{eqnarray}
This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
irregularities before the system equilibrates to a constant value.

>>>>

\subsection{Velocity Verlet}
The simulation is taking place in a NVE ensemble from a thermodynamic point of view. This means that the particle number $N$, the volume $V$ and the
total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
they have been drawn.\\
The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
of all the atoms in the system have to be added to get the resulting velocity vector for the whole system that is thought to be acting on the center
of mass. This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
velcocity vector of every atom in the system. 
    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
The angular momentum can be defined via the moment of inertia $\mathbf{I}$
    \label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
    \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
    \label{eq:angularmomentum2}
    \mathbf{L} = \mathbf{r} \times \mathbf{p}
where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
$\mathbf{I}$ for a filled cube is given by 
    \mathbf{I} = 
    \begin{pmatrix}
        \frac{mL^2}{6} & 0 & 0\\
        0 & \frac{mL^2}{6} & 0\\
        0 & 0 & \frac{mL^2}{6} 
    \end{pmatrix}
where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
equation~\eqref{eq:angularmomentum} and rewrite it as
    \label{eq:omega}
    \boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
between the particle and the center of mass
    \Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
    \mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. If the system should have a
specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
    E &=& \sum_{i=1}^N \mathbf{v}_i^2\\
    \lambda &=& \sqrt{\frac{3(N-1)T}{E}}\\
    \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i 
This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
irregularities before the system equilibrates to a constant value.
